Improving the Image Acquisition Rate of an Atomic Force Microscope Through Compressive Sampling

### 2019-05-28 latex/afm-cs-main.tex
Streamline latex. svg-inkscape should not be part of repo.

\documentclass[twocolumn,twoside]{IEEEtran/IEEEtran}
% \documentclass[defaultstyle,11pt]{IEEETran}
% Lucy says to look at:
% float_page_fraction
% tex_page_fraction
% https://tex.stackexchange.com/questions/35125/how-to-use-the-placement-options-t-h-with-figures#35130
% \renewcommand\floatpagefraction{0.5} %% default: 0.5
% \renewcommand\topfraction{0.7} % default: 0.7
% \renewcommand*\textfraction{.05}
\usepackage{amssymb}		% to get all AMS symbols
\usepackage{amsthm}
\usepackage{mathtools}
\usepackage{graphicx}		% to insert figures
\usepackage{hyperref}		% PDF hyperreferences??
\usepackage{multirow}
\usepackage{mathtools}
\usepackage{cite}
\usepackage{listings}
\usepackage{booktabs}
% \usepackage{amsthm}
\usepackage[percent]{overpic}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{xspace}
\usepackage{enumitem}
\usepackage{catchfilebetweentags} % for data input
\usepackage[inkscapelatex=false, inkscapepath=svgsubpath]{svg}
\usepackage{url}
\usepackage{mdframed}
\usepackage{amsmath, dsfont, bbold}
% \usepackage{algpseudocode,algorithm,algorithmicx}
\usepackage[ruled,linesnumbered]{algorithm2e}
\usepackage{placeins}
\usepackage{todonotes}

@@

\usepackage{layouts}
%textwidth: \printinunitsof{in}\prntlen{\textwidth}
%linewidth: \printinunitsof{in}\prntlen{\linewidth}
% textwidth: \printinunitsof{in}\prntlen{\textheight}

% this reveals that textwidth and linewidth are 6.5 inches,
% and textheight is 9.1 inches.

@@

\maketitle
\begin{abstract}
  Undersampling-based approaches in Atomic Force Microscopy (AFM) aim to reduce
  the time to acquire an image by reducing the number of measurements needed
  while still maintaining image quality. In this paper, we describe a hardware
  implementation and demonstration of one such approach based on the use of
  collections of short scans known as $\mu$--paths. Using a commercial AFM, we
  acquire data on two different grating samples using two different sampling
  approaches: randomly-placed $\mu-$paths and $\mu-$paths designed to minimize
  the reconstruction error based on partial prior information about the spatial
  frequency content in the sample image. Reconstructions are made from these
  data using three approaches: (1) inpainting, an interpolation technique that
  diffuses information from sampled locations into unsampled regions, (2)
  basis-pursuit, a compressive sensing-based algorithm that seeks to minimize a
  measure of sparsity in the underlying image, and (3) a new variant of basis
  pursuit, Basis Pursuit Vertical Variation (BPVV), that is designed to reduce
  artifacts arising from the sampling pattern. The quality of the resulting
  images is compared to images from standard raster scans of the same regions at
  comparable imaging rates using both the peak signal-to-noise ratio and the
  structured similarity index metric. These experiments demonstrate that at slow
  scan rates, the $\mu-$path scheme produces similar image quality with
  significantly less sampling (and thus less tip-sample interaction) and that at
  high rates, the undersampling scheme produces higher quality images than
  raster scanning.
\end{abstract}

\section{Introduction}\label{sec:introduction}
The Atomic Force Microscope (AFM) is a powerful instrument capable of imaging
sample surface topography, material characteristics, and other surface
properties, at the nanometer scale. AFMs acquire information about the sample
through a variety of imaging modes, all of which rely on the deflection of the
cantilever caused by the tip-sample interaction force. While the details of each
of the modes are quite different, in general, feedback control is used to hold
the deflection signal constant and information about the property of interest is
inferred from the applied control or the measured dynamics of the cantilever
\cite{Abramovitch:2007gt}. Because of its versatility, spatial resolution, and
ability to image in vacuum, air, and liquid, AFM is widely used in a variety of
disciplines, including physics, biology and materials science
\cite{Dufrene:2017gm,Yang:2017im,Payton:2016hj,Altman:2015ic,Haase:2015eh}

@@

In AFM, implementing a truly random pattern requires that the tip be engaged
with the surface to collect a measurement, lifted and moved to the new location,
and engaged again. Because the re-engagement process is typically slow, this can
lead to excessively long image acquisition times and negate the gains from
undersampling \cite{andersson2012non}. One discrete undersampling scheme, called
a $\mu$-path pattern, was proposed in \cite{maxwell2014compressed} to improve
sampling efficiency. The $\mu$-path pattern consists of randomly placed short,
horizontal scans (see Fig.~\ref{fig:mu_mask} for one such pattern) and is
designed to balance the randomness needed to ensure good reconstruction with
continuous scanning to reduce the number of tip
engagements. %Though the pattern is motivated from CS, it also works well in practice for inpainting reconstruction \cite{luo2015comparison}.

Our previous work on the $\mu$-path pattern using theoretical calculations,
simulations, and a preliminary implementation, demonstrates that the approach
can achieve significant scanning time reduction while maintaining faithful image
reconstruction \cite{maxwell2014compressed,Luo:2015tu,Braker:2018gi}. This paper
builds upon that earlier work in several significant ways. First, we demonstrate
performance using both randomly-placed $\mu-$paths and designed $\mu-$paths that
are optimized to minimize the expected reconstruction error given some initial
estimate of the frequency structure of the sample. These schemes are described
in Sec. \ref{sec:samplingScheme}. We also introduce a new reconstruction
algorithm which is designed to minimize the artifacts arising from the structure
of the $\mu$-path scanning pattern. This algorithm is described and compared to
existing reconstruction methods through simulations in Sec.
\ref{sec:reconstructionMethods}. We then demonstrate the entire scheme from
sampling through reconstruction using experiments on a commercial AFM (Agilent
5500). The experimental setup and implementation details are described in Secs.
\ref{sec:experimentalSetup} and \ref{sec:implementation}\footnote{The software
  portion of the implementation and image reconstruction described in this paper
  can be found at \url{https://github.com/yufan88/}}. Experimental results with
two different grating samples are provided in Sec.
\ref{sec:experimentalResults}. Finally, we provide a few concluding remarks in
Sec. \ref{sec:conclusions}.

% =========================================================================
\subsection{Reconstruction methods}
\label{sec:reconstructionMethods}

In this work, we use three reconstruction algorithms to generate the final
surface image. The first one comes from the field of compressive sensing (CS).
CS is a signal processing technique which aims at signal reconstruction from a
relatively small (sub-Nyquist limit) number of measurements
\cite{carmi2014compressive}. It takes advantage of the approximate sparsity of
real-world signals, that is, of the fact that many coefficients of such signals
are close to zero when represented in an appropriate basis. CS methods seek the
true image signal $x\in\mathbb{R}^n$ from the following observation equation,
\begin{equation}\label{op:observation}
  y = R x = RU\eta,
\end{equation}
\noindent where $y\in\mathbb{R}^n$ is the observation vector, $R$ is an
$n\times m$ matrix defining the measurements, $U$ is an $m\times m$ sparsity
basis and $\eta$ is the sparse representation of $x$ in the domain of $U$. In
general, $n\ll m$. In an imaging application, where the underlying signal is a
matrix $X\in\mathbb{R}^{h\times s}$, we take $x=\text{vec}(X)$, where the
$\text{vec}(\cdot)$ operator stacks the columns of a matrix.

In the AFM application, the probe can only measure a single pixel at a time.
Thus the rows of $R$ are a subset of the rows of an $m$ by $m$ identity matrix.
Ideally, the sparsity basis and
the measurement matrix $R$ will have a low \textit{mutual coherence}, a
measure that describes how each of the rows of $R$ (the measurements)
``spreads out'' in the domain of $U$ \cite{candes2007sparsity}. In practice,
the Discrete Cosine Transform (DCT) keeps a good balance between achieving a low
mutual coherence between the basis and the required structure of the AFM sensing
matrices and providing a high sparsity of typical AFM sample images. In the following,
we assume that $U^TU=I$, though in the simulation and experimental results, we take
$U$ as the DCT basis.

@@

Although BPDN is effective in the general setting, using it for reconstruction
from horizontal $\mu$-path samples often yields artifacts in the vertical
direction (that is, the direction orthogonal to the $\mu$-path scans) leading to
strong discontinuities in the image. These artifacts grow more prominent as the
length of the $\mu$-paths is increased \cite{maxwell2014compressed} as these
longer paths lead to increased mutual coherence between the measurement matrix
$R$ and the sparsity basis $U$. However, longer $\mu$-paths means fewer
tip re-engagements and thus shorter scan times.
As a result, the second
algorithm we consider is a new variant of BPDN that we term basis pursuit with
vertical variation (BPVV). In order to mitigate the vertical artifacts, BPVV
adds a vertical total variation penalty in the spatial domain to the
optimization objective. That is, \eqref{op:bp} is modified to

\begin{align}
  \min_{x\in Q_p}~f(x) &= \min_{x\in Q_p} |U^Tx|_1 + \alpha_v|D_vx|_1 + \alpha_h|D_hx|_1  \label{op:bp_tv}
  % \text{subject to} &\, \| R x - y \|_2 < \sigma,
\end{align}
where ${D_v x=\text{vec}(\nabla_v X)}$ (resp., ${D_v x=\text{vec}(\nabla_v X)}$) is the
discrete gradient in the vertical (resp., horizontal) direction and
\begin{align*}
  (\nabla_v X)_{i,j} = 
  \begin{cases}
    X_{i+1,j} - X_{i,j}, & 1\leq i<h,~ 1\leq j\leq s,\\
    0, & \phantom{:}i=h,~\phantom{:::::} 1\leq j\leq s.
  \end{cases}\\
  (\nabla_h X)_{i,j} = 
  \begin{cases}
    X_{i,j+1} - X_{i,j}, & 1\leq i\leq h,~1\leq j<s \\
    0, & \phantom{}1\leq i\leq,~\phantom{:::} j=s
  \end{cases}
\end{align*}
\todo{This is way too long}
In \eqref{op:bp_tv}, $\left \| D_v x \right \|_1$ (resp., $\left \| D_h x \right \|_1$) is the total variation (TV) of the signal in the vertical (resp., horizontal) direction. The parameters $\alpha_v$ and $\alpha_h$ are weighting parameters.
The description of BPVV in \eqref{op:bp_tv} can be viewed as the combination of CS
and inpainting (described below), in which the first term seeks a sparse solution while the
last two terms try to diffuse pixel information.
To solve the optimization problem in \eqref{op:bp_tv} we use a (slight) extension to NESTA~\cite{becker_nesta_2011}. To that end, we re-write the cost $f(x)$ in \eqref{op:bp_tv} as
\begin{equation*}
  f(x) = \max_{u\in Q_d} \left\langle u, Wx \right\rangle,\quad Q_d = \{u:~||u||_{\infty} \leq 1\}
\end{equation*}
where 
\begin{equation*}
W = \begin{bmatrix}U&\alpha_hD^T_h&\alpha_vD^T_v \end{bmatrix}^T.
%   W = \begin{bmatrix}U^T\\\alpha_hD_h\\\alpha_vD_v \end{bmatrix}.
\end{equation*}
One of the core features of the NESTA algorithm is to replace the non-smooth functional $f(x)$ with a smoothed approximation
\begin{equation}
  f_{\mu} = \max_{u\in Q_d} \left\langle u,
    W x \right\rangle - \frac{\mu}{2}||u||_2^2. \label{eqn:fcn_prox}
\end{equation}
The gradient of \eqref{eqn:fcn_prox} is given by
\begin{equation*}
  \nabla f_{\mu}(x)[i] = W^Tu_{\mu}(x)
\end{equation*}
where
\begin{equation*}
  (u_{\mu}(x))[i] = 
\begin{cases}
    \mu^{-1}(Wx)[i], & |W x[i]| < \mu,\\
    \text{sgn}(W x[i]), & \text{else}.
  \end{cases}
\end{equation*}

\begin{algorithm}
  \SetKwInOut{Initialize}{Initialize} \SetKwInOut{Output}{Output}

  \Initialize{$x \in Q_p, \mu=0.9||W x_0||_{\infty}$, $L_{\mu} = \frac{1}{\mu}||W||^2_2$}
    % $x^{k+1} = x^{k}$\\
  \While{$\mu < \mu_{\textrm{final}}$ } {
    $k=0$\\
      \While{$\Delta f_{\mu} > \delta$} {
        $\alpha_k = \frac{1}{2}(k+1)$ \\
        $\tau_k = \frac{2}{k+3}$ \\
        $\displaystyle y_k = \text{arg } \min_{x\in Q_p}
          \frac{L_{\mu}}{2}||x - x_k||_2^2 + \langle \nabla f(x_k),~ x\rangle$\\
        $\displaystyle {z_k = \text{arg } \min_{x\in Q_p} \frac{L_{\mu}}{2}||x-x_0||_2^2 +
        \langle \sum_{i=0}^k \alpha_i \nabla f(x_i), ~x \rangle}$\\
        $x_{k+1} = \tau_kz_k + (1-\tau_k)y_k$\\
        $k \leftarrow k + 1$\\
      }
      $x_0 = x_k$\\
      $\mu = \gamma \mu$\\
      $L_{\mu} = (1/\mu)||W||^2_2$
    }
  \Output{Reconstruction $x_k$}
  \caption{NESTA algorithm}
  \label{ag:BPTV}
\end{algorithm}
In the termination condition for the inner loop, $\Delta f_{\mu} = |f_{\mu}(x_k) - \bar{f}_{\mu}(x_k)|/\bar{f}_{\mu}(x_k)$ where
\begin{equation}
\bar{f}_{\mu}(x_k) = \frac{1}{\min\{10, k\}} \sum_{\ell=1}^{\min\{10,k\}} f_{\mu}(x_{k-\ell})
\end{equation}
The sub-problems in lines 6 and 7 have the form
\begin{equation}
  p = \textrm{arg } \min_{x\in Q_p} \frac{L}{2}||x - r|| + \langle g,~x - r\rangle
\end{equation}
for some vectors $g$ and $r$ and, as shown in~\cite{becker_nesta_2011}, have a closed-form solution given by
\begin{align}
  \lambda & = \textrm{max}(0, \sigma^{-1}\left| b - Rq\right| - L\\
  q &= r - L^{-1}g\\
  p &= \left(I - \frac{\lambda}{\lambda + L} R^TR\right)\left( \frac{\lambda}{L}R^Tb + q\right).
\end{align}

@@

A comparison between BPDN and BPVV is shown in
Fig.~\ref{fig:BPTV_demonstration}. Images of three different samples (a circular
grating, a square grating and a DNA image) were acquired using an Agilent 5500
AFM with a standard raster scan (first and fourth columns in the figure). These
images were then resampled using a random $\mu$-path pattern of size 40 pixels
with 25\% of the pixels sampled. Images were reconstructed using BPDN (second
and fifth columns) and BPVV (third and sixth columns). The reconstruction
results show the artifacts apparent in the BPDN reconstruction are largely
mitigated using BPVV.

% =========================================================================
\section{Experimental setup} \label{sec:experimentalSetup}
% =========================================================================
Our experimental setup, illustrated in Fig.~\ref{fig:exp_setup_anderssonLab},
consists of an Agilent 5400 AFM, a cRIO-9082 embedded controller (National
Instruments), and a standard desktop computer. Through a breakout box, the
Agilent 5500 provides access to the $x$, $y$, and $z$-axis measurements as well
as the deflection signal (denoted in the following as $l$). When the Agilent
default software (PicoView) is set to open-loop mode, a $\pm 10$ V input on the
control box allows control of the $x$, $y$, and $z$-axis piezos. The N9524A
piezoelectric scanner in the Agilent 5400 provides a maximum scan area of
$90~\mu$m $\times$ 90~$\mu$m in the $xy$ (horizontal) direction and a range of
$20~\mu$m in the $z$ (vertical) direction. Initial engagement of the tip to the
sample is performed using the PicoView software before control is handed over to
the custom controllers.

\todo{ add stuff about nPoint stage and C300.}

All control logic is programmed using LabVIEW 2017 and compiled to a Xilinx
Spartan-6 LX150 Field Programmable Gate Array (FPGA) inside the cRIO-9082. The
cRIO includes a 16-Bit, 100 kHz NI-9215 analog-to-digital input module and a
16-Bit, 100 kHz NI-9263 digital-to-analog output module. All control loops are
implemented using a 25 kHz sampling rate.

% =========================================================================
\section{Control}\label{sec:control}
The control structure for each axis is different. For the $X$-axis, we use a
state space controller. It is described fully in \cite{braker_afmmpc_2019} and
achieves a bandwidth of about 110 Hz, which is approximately 1/3 of the $X$-axis
resonance at 350 Hz.

The control law used for the $Y$-axis is a simple integral controllers of the
form
\begin{equation}
  \frac{U(z)}{E_Y(z)}=D(z) = \frac{K_Iz}{z-1},
  \label{eqn:dzI}
\end{equation}
where, the error signal $E_Y(z) = Y(z)-R_Y$.

@@

>>>>

\documentclass[twocolumn,oneside]{IEEEtran/IEEEtran}
% \documentclass[defaultstyle,11pt]{IEEETran}
% Lucy says to look at:
% float_page_fraction
% tex_page_fraction
% https://tex.stackexchange.com/questions/35125/how-to-use-the-placement-options-t-h-with-figures#35130
% \renewcommand\floatpagefraction{0.5} %% default: 0.5
% \renewcommand\topfraction{0.7} % default: 0.7
% \renewcommand*\textfraction{.05}
\usepackage{showframe}
\usepackage{amssymb}		% to get all AMS symbols
\usepackage{amsthm}
\usepackage{mathtools}
\usepackage{graphicx}		% to insert figures
\usepackage{hyperref}		% PDF hyperreferences??
\usepackage{multirow}
\usepackage{mathtools}
\usepackage{cite}
\usepackage{listings}
\usepackage{booktabs}
% \usepackage{amsthm}
\usepackage[percent]{overpic}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{xspace}
\usepackage{enumitem}
\usepackage{catchfilebetweentags} % for data input
\usepackage[inkscapelatex=false, inkscapepath=svgsubpath]{svg}
\usepackage{url}
\usepackage{mdframed}
\usepackage{amsmath, dsfont, bbold}
% \usepackage{algpseudocode,algorithm,algorithmicx}
\usepackage[ruled,linesnumbered]{algorithm2e}
\usepackage{placeins}
\usepackage{todonotes}

@@

\newcommand{\rzup}{\ensuremath{r_{z,\textrm{up}}}\xspace}
\newcommand{\rzs}{\ensuremath{r_{z,\textrm{s}}}\xspace}

% *******************************************************************************
% This artificially increases the papersize and margins, so todo notes are legible.
% ***************** WARNING: comment this out for final version. ****************
\usepackage[paperwidth=734.295pt, textwidth=516pt,
            marginparsep=5pt, hoffset=10pt, marginparwidth=90pt,
            textheight=696pt]{geometry}
% Use layouts to display the original dimensions, so we can get as close to that as possible with
% our fake page size.
\usepackage{layouts}
%textwidth: \printinunitsof{in}\prntlen{\textwidth}
%linewidth: \printinunitsof{in}\prntlen{\linewidth}
% textwidth: \printinunitsof{in}\prntlen{\textheight}

% *******************************************************************************

@@

\maketitle
\begin{abstract}
  Undersampling-based approaches in Atomic Force Microscopy (AFM) aim to reduce
  the time to acquire an image by reducing the number of measurements needed
  while still maintaining image quality. In this paper, we describe a hardware
  implementation and demonstration of one such approach based on the use of
  collections of short scans known as $\mu$--paths. Using a commercial AFM, we
  acquire data on a grating sample.
  %using two different sampling
  %approaches: randomly-placed $\mu-$paths and $\mu-$paths designed to minimize
  %the reconstruction error based on partial prior information about the spatial
  %frequency content in the sample image.
  Reconstructions are made from these data using
  %three approaches: (1) inpainting, an interpolation technique that
  %diffuses information from sampled locations into unsampled regions, (2)
  %basis-pursuit, a compressive sensing-based algorithm that seeks to minimize a
  %measure of sparsity in the underlying image, and (3)
  a new variant of basis
  pursuit, Basis Pursuit with Vertical Variation (BPVV), that is designed to reduce
  artifacts arising from the sampling pattern. The quality of the resulting
  images is compared to images from standard raster scans of the same regions at
  comparable imaging rates using both the peak signal-to-noise ratio and the
  structured similarity index metric
  and a new metric we call the relative damage index.
  \todo{Will this conclusion still hold?}
  These experiments demonstrate that at slow
  scan rates, the $\mu-$path scheme produces similar image quality with
  significantly less sampling (and thus less tip-sample interaction) and that at
  high rates, the undersampling scheme produces higher quality images than
  raster scanning.
\end{abstract}

\section{Introduction}\label{sec:introduction}
% textwidth: \printinunitsof{pt}\prntlen{\textwidth}

% linewidth: \printinunitsof{pt}\prntlen{\linewidth}

% paperwidth: \printinunitsof{pt}\prntlen{\paperwidth}

% marginparwidth: \printinunitsof{pt}\prntlen{\marginparwidth}

% marginsep: \printinunitsof{pt}\prntlen{\marginparsep}

% textheight: \printinunitsof{pt}\prntlen{\textheight}

The Atomic Force Microscope (AFM) is a powerful instrument capable of imaging
sample surface topography, material characteristics, and other surface
properties, at the nanometer scale. AFMs acquire information about the sample
through a variety of imaging modes, all of which rely on the deflection of the
cantilever caused by the tip-sample interaction force. While the details of each
of the modes are quite different, in general, feedback control is used to hold
the deflection signal constant and information about the property of interest is
inferred from the applied control or the measured dynamics of the cantilever
\cite{Abramovitch:2007gt}. Because of its versatility, spatial resolution, and
ability to image in vacuum, air, and liquid, AFM is widely used in a variety of
disciplines, including physics, biology and materials science
\cite{Dufrene:2017gm,Yang:2017im,Payton:2016hj,Altman:2015ic,Haase:2015eh}

@@

In AFM, implementing a truly random pattern requires that the tip be engaged
with the surface to collect a measurement, lifted and moved to the new location,
and engaged again. Because the re-engagement process is typically slow, this can
lead to excessively long image acquisition times and negate the gains from
undersampling \cite{andersson2012non}. One discrete undersampling scheme, called
a $\mu$-path pattern, was proposed in \cite{maxwell2014compressed} to improve
sampling efficiency. The $\mu$-path pattern consists of randomly placed short,
horizontal scans (see Fig.~\ref{fig:mu_mask} for one such pattern) and is
designed to balance the randomness needed to ensure good reconstruction with
continuous scanning to reduce the number of tip
engagements. %Though the pattern is motivated from CS, it also works well in practice for inpainting reconstruction \cite{luo2015comparison}.
\todo{Would be beneficial to shrink intro, if possible}

Our previous work on the $\mu$-path pattern using theoretical calculations,
simulations, and a preliminary implementation, demonstrates that the approach
can achieve significant scanning time reduction while maintaining faithful image
reconstruction \cite{maxwell2014compressed,Luo:2015tu, braker_hardware_2018}. This paper
builds upon that earlier work in several significant ways.
% % ------------------ TKTKTKT ------------------
% \todo{Strike this?}
% First, we demonstrate
% performance using both randomly-placed $\mu-$paths and designed $\mu-$paths that
% are optimized to minimize the expected reconstruction error given some initial
% estimate of the frequency structure of the sample. These schemes are described
% in Sec. \ref{sec:samplingScheme}.
% % --------------------------
We also introduce a new reconstruction
algorithm which is designed to minimize the artifacts arising from the structure
of the $\mu$-path scanning pattern. This algorithm is described and compared to
existing reconstruction methods through simulations in Sec.
\ref{sec:reconstructionMethods}.
Section \ref{sec:control} describes improvements to the control algorithms compared to our preliminary implementation in ~\cite{braker_hardware_2018}.
We then demonstrate the entire scheme from
sampling through reconstruction using experiments on a commercial AFM (Agilent
5500). The experimental setup and implementation details are described in Secs.
\ref{sec:experimentalSetup} and \ref{sec:implementation}\footnote{The software
  portion of the implementation and image reconstruction described in this paper
  can be found at \url{https://github.com/rabraker/AFM_CS}}. Experimental results are provided in Sec.
\ref{sec:results}. Finally, we provide a few concluding remarks in
Sec. \ref{sec:conclusions}.

\begin{figure}
  \centering
  \includegraphics[width=0.7\columnwidth]{figures-SBA/random_mask.pdf}
  \caption{An example of the horizontal $\mu$-path sampling pattern for a 256$\times$256 pixel image with 35 pixels in each short scan and a total of 20\% pixels scanned.}
  \label{fig:mu_mask}
\end{figure}
% =========================================================================
\subsection{Reconstruction methods}
\label{sec:reconstructionMethods}
Compressive sensing (CS) is a signal processing technique which aims at signal reconstruction from a
relatively small (sub-Nyquist limit) number of measurements
\cite{carmi2014compressive}. It takes advantage of the approximate sparsity of
real-world signals, that is, that many coefficients of such signals
are close to zero when represented in an appropriate basis. CS methods seek the
true image signal $x\in\mathbb{R}^n$ from the observation equation,
\begin{equation}\label{op:observation}
  y = R x = RU\eta,
\end{equation}
\noindent where $y\in\mathbb{R}^n$ is the observation vector, $R$ is an
$n\times m$ matrix defining the measurements, $U$ is an $m\times m$ sparsity
basis and $\eta$ is the sparse representation of $x$ in the domain of $U$. In
general, $n\ll m$. In an imaging application, where the underlying signal is a
matrix $X\in\mathbb{R}^{h\times s}$, we take $x=\text{vec}(X)$, where the
$\text{vec}(\cdot)$ operator stacks the columns of a matrix.

In the AFM application, the probe can only measure a single pixel at a time.
Thus the rows of $R$ are a subset of the rows of an $m$ by $m$ identity matrix.
Ideally, the sparsity basis and
the measurement matrix $R$ will have a low \textit{mutual coherence}, a
measure that describes how each of the rows of $R$ (the measurements)
``spreads out'' in the domain of $U$ \cite{candes2007sparsity}.
In the following, we assume that $U^TU=I$, though in the simulation and
experimental results, we take $U$ as the Discrete Cosine Transform (DCT) basis.
The DCT basis generally keeps a good balance between achieving a low
mutual coherence between $U$ and the required structure of the AFM sensing
matrices $R$ and providing a high sparsity of typical AFM sample images.

@@

Although BPDN is effective in the general setting, using it for reconstruction
from horizontal $\mu$-path samples often yields artifacts in the vertical
direction (that is, the direction orthogonal to the $\mu$-path scans) leading to
strong discontinuities in the image. These artifacts grow more prominent as the
length of the $\mu$-paths is increased \cite{maxwell2014compressed} as these
longer paths lead to increased mutual coherence between the measurement matrix
$R$ and the sparsity basis $U$. However, longer $\mu$-paths means fewer
tip re-engagements and thus shorter scan times.

\todo{This development is way too long!}
To mitigate these affects, we propose a new variant of BPDN that we term basis pursuit with
vertical variation (BPVV). BPVV adds a vertical total variation penalty in the spatial domain to the
optimization objective. That is, \eqref{op:bp} is modified to
\todo{It has been my observation that some benefit is gained by including $D_h$, but weighting it less than $D_v$. Not sure if we should leave it in. If so, BPVV should be BPTV or something.}
\begin{align}
  \min_{x\in Q_p}~f(x) &= \min_{x\in Q_p} \left\|U^Tx\right\|_1 + \alpha_v\left\|D_vx\right\|_1
                         + \alpha_h\left\|D_hx\right\|_1  \label{op:bp_tv}
  % \text{subject to} &\, \| R x - y \|_2 < \sigma,
\end{align}
where ${D_v x=\text{vec}(\nabla_v X)}$ (resp., ${D_v x=\text{vec}(\nabla_v X)}$) is the
discrete gradient in the vertical (resp., horizontal) direction (see, e.g., \cite[Section 6.1,]{becker_nesta_2011}). 
% and
% \begin{align*}
%   (\nabla_v X)_{i,j} = 
%   \begin{cases}
%     X_{i+1,j} - X_{i,j}, & 1\leq i<h,~ 1\leq j\leq s,\\
%     0, & \phantom{:}i=h,~\phantom{:::::} 1\leq j\leq s.
%   \end{cases}\\
%   (\nabla_h X)_{i,j} = 
%   \begin{cases}
%     X_{i,j+1} - X_{i,j}, & 1\leq i\leq h,~1\leq j<s \\
%     0, & \phantom{}1\leq i\leq,~\phantom{:::} j=s
%   \end{cases}
% \end{align*}
Thus, in \eqref{op:bp_tv}, $\left \| D_v x \right \|_1$ (resp., $\left \| D_h x \right \|_1$) is the total variation (TV) of the signal in the vertical (resp., horizontal) direction. The parameters $\alpha_v$ and $\alpha_h$ are weighting parameters.

The description of BPVV in \eqref{op:bp_tv} can be viewed as the combination of CS
and inpainting, in which the first term seeks a sparse solution while the
last two terms try to diffuse pixel information.
\todo{When written this way, it seems that BP+VV/TV can could also be interpreted in the frameowrk of overcomplete dictionaries?\cite{candes_redundant_2011}. }
To solve the optimization problem in \eqref{op:bp_tv} we use a (slight) extension to NESTA~\cite{becker_nesta_2011}. To that end, we re-write the cost $f(x)$ in \eqref{op:bp_tv} as
\begin{equation*}
  f(x) = \max_{u\in Q_d} \left\langle u, Wx \right\rangle,\quad Q_d = \{u:~||u||_{\infty} \leq 1\}
\end{equation*}
where
\begin{equation*}
W = \begin{bmatrix}U&\alpha_hD^T_h&\alpha_vD^T_v \end{bmatrix}^T.
%   W = \begin{bmatrix}U^T\\\alpha_hD_h\\\alpha_vD_v \end{bmatrix}.
\end{equation*}
One of the core features of the NESTA algorithm is to replace the non-smooth functional $f(x)$ with a smoothed approximation
\begin{equation}
  f_{\mu} = \max_{u\in Q_d} \left\langle u,
    W x \right\rangle - \frac{\mu}{2}||u||_2^2. \label{eqn:fcn_prox}
\end{equation}
The gradient of \eqref{eqn:fcn_prox} is given by
\begin{equation*}
  \nabla f_{\mu}(x)[i] = W^Tu_{\mu}(x)
\end{equation*}
where
\begin{equation*}
  (u_{\mu}(x))[i] = 
\begin{cases}
    \mu^{-1}(Wx)[i], & |W x[i]| < \mu,\\
    \text{sgn}(W x[i]), & \text{else}.
  \end{cases}
\end{equation*}
The optimization \eqref{op:bp_tv} may now be solved using Algorithm~\ref{ag:NESTA}.
\todo{seems like a waste of space to reprint the algo, since its really no different than the reference}
\begin{algorithm}
  \SetKwInOut{Initialize}{Initialize} \SetKwInOut{Output}{Output}

  \Initialize{$x \in Q_p, \mu=0.9||W x_0||_{\infty}$, $L_{\mu} = \frac{1}{\mu}||W||^2_2$}
    % $x^{k+1} = x^{k}$\\
  \While{$\mu < \mu_{\textrm{final}}$ } {
    $k=0$\\
      \While{$\Delta f_{\mu} > \delta$} {
        $\alpha_k = \frac{1}{2}(k+1)$ \\
        $\tau_k = \frac{2}{k+3}$ \\
        $\displaystyle y_k = \text{arg } \min_{x\in Q_p}
          \frac{L_{\mu}}{2}||x - x_k||_2^2 + \langle \nabla f(x_k),~ x\rangle$\\
        $\displaystyle {z_k = \text{arg } \min_{x\in Q_p} \frac{L_{\mu}}{2}||x-x_0||_2^2 +
        \langle \sum_{i=0}^k \alpha_i \nabla f(x_i), ~x \rangle}$\\
        $x_{k+1} = \tau_kz_k + (1-\tau_k)y_k$\\
        $k \leftarrow k + 1$\\
      }
      $x_0 = x_k$\\
      $\mu = \gamma \mu$\\
      $L_{\mu} = (1/\mu)||W||^2_2$
    }
  \Output{Reconstruction $x_k$}
  \caption{NESTA algorithm}
  \label{ag:NESTA}
\end{algorithm}

In the termination condition for the inner loop, $\Delta f_{\mu} = |f_{\mu}(x_k) - \bar{f}_{\mu}(x_k)|/\bar{f}_{\mu}(x_k)$ where
\begin{equation}
\bar{f}_{\mu}(x_k) = \frac{1}{\min\{10, k\}} \sum_{\ell=1}^{\min\{10,k\}} f_{\mu}(x_{k-\ell})
\end{equation}
When $RR^T=I$, the sub-problems in lines 6 and 7 have a closed-form solution. Details can be found in~\cite{becker_nesta_2011}.

% The sub-problems in lines 6 and 7 have the form
% \begin{equation}
%   p = \textrm{arg } \min_{x\in Q_p} \frac{L}{2}||x - r|| + \langle g,~x - r\rangle
% \end{equation}
% for some vectors $g$ and $r$ and, as shown in~\cite{becker_nesta_2011}, have a closed-form solution given by
% \begin{align}
%   \lambda & = \textrm{max}(0, \sigma^{-1}\left| b - Rq\right| - L\\
%   q &= r - L^{-1}g\\
%   p &= \left(I - \frac{\lambda}{\lambda + L} R^TR\right)\left( \frac{\lambda}{L}R^Tb + q\right).
% \end{align}

@@

A comparison between BPDN and BPVV is shown in
Fig.~\ref{fig:BPTV_demonstration}. Images of two different samples
(a square grating and a DNA image) were acquired using an Agilent 5500
AFM with a standard raster scan (first and fourth columns in the figure). These
images were then resampled using a random $\mu$-path pattern of size 40 pixels
with 25\% of the pixels sampled. Images were reconstructed using BPDN (second
and fifth columns) and BPVV (third and sixth columns). The reconstruction
results show the artifacts apparent in the BPDN reconstruction are largely
mitigated using BPVV. Thus, in the remainder of the paper, we focus exclusively on BPVV.

% =========================================================================
\section{Experimental setup} \label{sec:experimentalSetup}
% =========================================================================
Our experimental setup, illustrated in Fig.~\ref{fig:exp_setup_anderssonLab},
consists of an Agilent 5400 AFM  retrofitted with an nPoint NPXY100A
piezoelectric stage, a cRIO-9082 embedded controller (National
Instruments), and a standard desktop computer.

Through a breakout box, the Agilent 5400 provides access to the $z$-axis deflection signal. When the Agilent software is set to open-loop mode, a $\pm10$~volt input on the standard control box allows control of the $z$-axis of the N9524a piezo scanner, which has a total range of 7 $\mu$m.
Initial (course) engagement of the tip to the
sample is performed using the PicoView software before control is handed over to
the custom controllers.

All control logic is programmed using LabVIEW 2019 and compiled to a Xilinx
Spartan-6 LX150 Field Programmable Gate Array (FPGA) inside the cRIO-9082. The
cRIO includes a 16-Bit, 100 kHz NI-9215 analog-to-digital input module and a
16-Bit, 100 kHz NI-9263 digital-to-analog output module. All control loops are
implemented using a 25 kHz sampling rate.

% =========================================================================
\section{Control}\label{sec:control}
The control structure for each axis is different. For the $X$-axis, we use a
state space controller. It is described fully in \cite{braker_afmmpc_2019} and
achieves a bandwidth of about 150 Hz, which is over 1/3 of the first $X$-axis
resonance at 350 Hz.

The control law used for the $Y$-axis is a simple PI controller of the
form
\begin{equation}
  \frac{U(z)}{E_Y(z)}=D(z) = \frac{K_Iz}{z-1},
  \label{eqn:dzI}
\end{equation}
where, the error signal $E_Y(z) = Y(z)-R_Y$.

@@

\begin{figure*}[t]
  \centering
  \includegraphics[width=0.75\textwidth]{figures/AFM_loop_Dinv_antiwup-crop.pdf}
  \caption{Block diagram of the $Z$-axis control loop. The anti-windup is active
    for decision index $i\in\{0,1,5\}$, which correspond to states such that
    $r_Z=\rzup$.}
  \label{fig:afm_bd_dinv}
\end{figure*}
